---
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=10, fig.height=8, warning=FALSE, message=FALSE)
```

# Statistical Inference Course Project 2

## Overview
Loading the ToothGrowth data to perform a exploratory data analyses:
1 Basic summary of the data.
2 Confidence intervals and/or hypothesis tests and compare tooth growth by supp and dose. 
3 State the conclusions and the assumptions. 


## Load Data

```{r}
# load neccesary libraries
library(ggplot2)
library(datasets)
library(gridExtra)
library(GGally)

# The Effect of Vitamin C on Tooth Growth in Guinea Pigs
data(ToothGrowth)
toothGrowth <- ToothGrowth 
toothGrowth$dose <- as.factor(toothGrowth$dose) # convert to factor
```


## Basic Summary of the data
```{r}
str(toothGrowth)
summary(toothGrowth)
head(toothGrowth)

table(toothGrowth$supp, toothGrowth$dose)
```
```{r echo=FALSE}
p1 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose), colour="#CC0000") +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~supp) 

p2 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot() + 
  theme(legend.position="none") + 
  facet_grid(.~dose) 

p3 <- ggplot(data=toothGrowth, aes(x=supp,y=len,fill=supp)) +
  geom_boxplot()

p4 <- ggplot(data=toothGrowth, aes(x=dose,y=len,fill=dose)) +
  geom_boxplot()

grid.arrange(p1, p4, p2, p3, ncol = 2, nrow=2)

#ggpairs(toothGrowth, 
#        upper = list(continuous = "density", combo = "box"),
#        lower = list(continuous="smooth"),        
#        colour = "dose",
#        axisLabels='show')
```

Analysis based on Analysis of Variance
```{r}
anova.out <- aov(len ~ supp * dose, data=toothGrowth)
summary(anova.out)
```

The results show an interaction between the length and dosage (F(1,54)=15.572;p<0.01)
Also a very clear effect on length by supplement type (F(2,54)=92;p<0.01). 
Last but not least there is a minor interaction between the combination of supplement type  and dosage compared to the length(F(2,54)=4.107;p<0.05).

```{r}
TukeyHSD(anova.out)
```

The Tukey  analysis showsthe differences between each of the groups in supp and dose
Only the interactions between VC:0.5-OJ:0.5; VC:1-OJ:0.5; OJ:2-OJ:1; VC:2-OJ:1 and VC:2-OJ:2  are not significant
```{r}
confint(anova.out)
print(model.tables(anova.out,"means"),digits=3)
```

## Conclusions

Indications that both the supplement as the dosage have indipendent effects on the length of teeth.
Supplement type has a clear influence, but OJ has a greater avarage teethgrowth in combination with dosages 0.5 and 1 then for the VC supplement, while teeth length for the VC supplement vs the OJ in combiantion with dosage 2 has no significant effect (almost same mean & same confidence interval)