---
output:
  html_document: null
  keep_md: yes
  pdf_document: default
---

```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.width=7, fig.height=4, warning=FALSE, message=FALSE)
```

# Statistical Inference Course Project 1

## Overview
This project I'm intenting to investigate the exponential distribution in R and compare it with the Central Limit Theorem. 
The exponential distribution will be simulate in R with rexp(n, lambda) where lambda, is rate parameter. 
The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. 
It will set lambda = 0.2 for all of the simulations. Also I'm going to  investigate the distribution of averages of 40 exponentials. 

## Simulations

```{r}
# load libraries
library(ggplot2)

# set constants
lambda <- 0.2 # lambda
n <- 40 # exponetials
numberOfSimulations <- 1000 # quant of tests

# set the seed to create reproducability
set.seed(100)

# run the test resulting in n x numberOfSimulations 
exponentialDistributions <- matrix(data=rexp(n * numberOfSimulations, lambda), nrow=numberOfSimulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))
```

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, fill='#56B4E9', color="darkred") +   
  scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))
```

## Sample Mean vs Theoretical Mean

Expected mean $\mu$ of a exponential distribution of rate $\lambda$ is 

$\mu= \frac{1}{\lambda}$ 

```{r}
mu <- 1/lambda
mu
```

Let $\bar X$ be the average sample mean of 1000 simulations of 40 randomly sampled exponential distributions.

```{r}
meanOfMeans <- mean(exponentialDistributionMeans$means)
meanOfMeans
```

You can see the expected mean and the avarage sample are close 


## Sample Variance vs Theoretical Variance

Expected standard deviation $\sigma$ of a exponential distribution of rate $\lambda$ is 

$\sigma = \frac{1/\lambda}{\sqrt{n}}$ 

The e

```{r}
sd <- 1/lambda/sqrt(n)
sd
```

The variance $Var$ of standard deviation $\sigma$ is

$Var = \sigma^2$ 

```{r}
Var <- sd^2
Var
```

Let $Var_x$ be the variance of the average sample mean of 1000 simulations of 40 randomly sampled exponential distribution, and $\sigma_x$ the corresponding standard deviation.
```{r}
sd_x <- sd(exponentialDistributionMeans$means)
sd_x

Var_x <- var(exponentialDistributionMeans$means)
Var_x
```

As you can see the standard deviations are very close
Since variance is the square of the standard deviations, minor differnces will we enhanced, but are still pretty close.

## Distribution

Comparing the population means & standard deviation with a normal distribution of the expected values. Added lines for the calculated and expected means

```{r echo=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
  geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
  stat_function(fun = dnorm, arg = list(mean = mu , sd = sd), colour = "red", size=1) + 
  geom_vline(xintercept = mu, size=1, colour="#CC0000") + 
  geom_density(colour="blue", size=1) +
  geom_vline(xintercept = meanOfMeans, size=1, colour="#56B4E9") + 
  scale_x_continuous(breaks=seq(mu-3,mu+3,1), limits=c(mu-3,mu+3)) 
```

The graph shows the calculated distribution of means of random sampled exponantial distributions, overlaps quite nice with the normal distribution with the expected values based on the given lamba